---
title: "Project_1"
author: "Victor Torres"
date: "2024-02-17"
output: html_document
---
# DESCRIPTION.#
#### The purpose of this project is to analize data from a TXT file with chess tournament results. Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents, will be the information required to analized the data, and obtain all the information required for the project.

```{r load require libraries}
library(tidyverse)
library(dplyr)
library(stringr)
```
```{r Get the TXT file from my GitHub account}
#get raw data from my GitHub Account
raw_data <- readLines("https://raw.githubusercontent.com/vitugo23/DATA607/main/project1/Chess_tournament.txt")
```
```{r Extract Key fields, and putting it into a data frame}
#extract key fields and putting into a dataframe using Regex
player_num <- as.numeric(unlist(str_extract_all(raw_data,"(?<=\\s{3,4})\\d{1,2}(?=\\s)")))
player_name <- unlist(str_extract_all(raw_data,"(?<=\\d\\s\\|\\s)([A-z, -]*\\s){1,}[[:alpha:]]*(?=\\s*\\|)"))
player_state <- unlist(str_extract_all(raw_data, "[[:upper:]]{2}(?=\\s\\|)"))
total_pts <- as.numeric(unlist(str_extract_all(raw_data, "(?<=\\|)\\d\\.\\d")))
player_pre_rat <- as.numeric(unlist(str_extract_all(raw_data, "(?<=R:\\s{1,2})(\\d{3,4}(?=\\s))|(\\d{3,4}(?=P\\d{1,2}\\s*-))")))
processed_data <- data.frame(player_num, player_name, player_state, total_pts, player_pre_rat)
```
```{r review dataframe structure }
#review dataframe structure
str(processed_data)
```
```{r create a secondary row with opponents to compare the data between players}
#create a secondary row with opponents to compare the data between players
secondary_rows <- raw_data[seq(5, 196, 3)]
opponent_num <- as.numeric(unlist(str_extract_all(secondary_rows, "(?<=\\|(W|L|D)\\s{2,3})[[:digit:]]{1,2}(?=\\|)|((?<!->)(?<=\\|(U|H|B|X))\\s{4}(?=\\|))")))
```
```{r Calculate the PCR (average pre chess rating) for opponents}
#Calculate the PCR (average pre chess rating) for opponents
pcr_matrix <- matrix(data = NA, nrow = 64, ncol = 2)

# Assign names 'total_opp_pcr, and avg_opp_pcr to dataframe 
colnames(pcr_matrix) <- c("total_opp_pcr", "avg_opp_pcr")
row_counter <- 0
for(i in seq(from=1, to=length(opponent_num)-6, by=7)){
  row_counter <- row_counter + 1
  
  pcr_matrix[row_counter, 1] <- (sum(subset(processed_data$player_pre_rat, processed_data$player_num %in% opponent_num[seq(from=i, to=i+6, by=1)])))
  
# Calculate the average score per row
  pcr_matrix[row_counter, 2] <- pcr_matrix[row_counter, 1] / length(subset(opponent_num[seq(from=i, to=i+6, by=1)],!is.na(opponent_num[seq(from=i, to=i+6, by=1)])))
  
}
# Verify matrix works as need it
head(pcr_matrix, 5)
```
```{r Round to the nearest whole number}
#Round to the nearest whole number
pcr_matrix[, 2] <- round(pcr_matrix[,2], digits = 0)

processed_data <- cbind(processed_data, pcr_matrix[, 2])
processed_data <- rename(processed_data, avg_opp_pcr = `pcr_matrix[, 2]`)
```
```{r}
# Get working directory path
path <- getwd()

# Export file to working directory
write.csv(processed_data, file.path(path, "chess_processed_data.csv"))
head(processed_data, 5)
```
```{r graphics}
library(ggplot2)
processed_data %>%
  ggplot(aes(x = total_pts)) + geom_bar()

barplot(rbind(processed_data$`player_pre_rat`[41:64], processed_data$`avg_opp_pcr`[41:64]), beside = TRUE, col = c("yellow", "green"), 
        xlab = "Player Pre-Rat. vs. Average Opponent Pre-Rat.", ylab = "Players pre-rating", names.arg = c(41:64)) 


```


# Conclussion. #
### The data was succesfully extracted from the text file, loaded into a data frame, and analized in order to get the information required in the project. The process was succesfully achieved thanks to some of the Tidyverse functions as well as Regex.###

## Sources.##
###### R for Data Science ######
###### Stack Overflow ######
###### GeeksforGeeks ######
###### DataCamp ######

