---
title: "Week7"
author: "Victor Torres"
date: "2024-03-05"
output: html_document
---
# Introduction. #
### Week seven assignment is about preparing three different files with same data in it, to practice loading diferent formats into R Studio (HTML, JSON and XML). After convert the files to data frames, we have to compare to see if there is any differences, in order to understand better the file structures, the three files were manually written and then loaded into Github for reference and reproducibility. The files must contain the following information:Title, Authors, and two or three other attributes such as Genre, rating, and publisher. ###
#### Load the required libraries for the project. ####
```{r load libraries}
library (tidyverse)
library(RCurl)
library(rvest)
library(rjson)
library(XML)
library(methods)
library(compare)
```
## Load HTML file and display data in console. ##
```{r}
html_file = getURL("https://raw.githubusercontent.com/vitugo23/DATA607/main/Assignments/W7_Assignment/Book_table.html")
books.html <- readHTMLTable(html_file)
books.html <-books.html[[1]]
books.html
```
## Load XML files and display data in console. ##
```{r}
xml_file = getURL("https://raw.githubusercontent.com/vitugo23/DATA607/main/Assignments/W7_Assignment/Book_table.xml")
 books.xml <- xmlToDataFrame(xml_file)
 books.xml
```


## Load JSON file and display data in console. ##
```{r}
json_file = getURL("https://raw.githubusercontent.com/vitugo23/DATA607/main/Assignments/W7_Assignment/Book_table.json")
books.json <- fromJSON(json_file)
books.json <- data.table::rbindlist(books.json)
books.json
```
## Make a quick comparison between the three data frames, using .STR and compare function. ##
```{r using str function}
str(books.html)
```
```{r use str function 2}
str(books.xml)
```
```{r  use str function 3}
str(books.json)
```
```{r use compare function}
compare(books.html,books.xml,equal=TRUE)
```
```{r use compare function2}
compare(books.html,books.json,equal=TRUE)
```
```{r use compare function3}
compare(books.json,books.xml,equal = TRUE)
```
# Conclussion #
### After making a quick comparison between the three files, we can see that the data types of it are the same with 3 observations, and 5 variables.###
### There is a difference in the JSON file, since the coding structure differs from the other two data types. ###




