---
title: "Assignment2"
author: "Victor Torres"
date: "2024-02-03"
---
##Assignment 2
In this assignment we were requested to create a database and uploaded to the R console. I created a simple database in MySQL 8 with two tables with names of the movies and viewers, including the rating of each one.
##First we need to install the RMySQl package into R studio.
```{r}

library(RMySQL)
library(tidyverse)
```

##We need to create an R object (mysqlconnection) to access to the database.
##with username, password, DBname, host, port,

```{r}
mysqlconnection = dbConnect(RMySQL::MySQL(),
                            dbname='movie_rating',
                            host='localhost',
                            port=3306,
                            user='root',
                            password='Viviana10!')
```
##this command is to display the tables in the database(movie_rating)

```{r}
dbListTables(mysqlconnection)
```
## I created an object (mov_dat) to access to the table information.
```{r}
mov_dat <- dbGetQuery(mysqlconnection, "select * from movies")
glimpse(mov_dat)
```

