---
title: "Week9_Assignment"
author: "Victor Torres"
date: "2024-03-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intoduction.

#### For this week's assignment, we need to choose one APIs from the New York Times web page, get an API key, construct an interface in R to read the JSON data, and and transform it into an R DataFrame. 

### Import Libraries required for this assignment.

```{r libraries}
library(tidyverse)
library(jsonlite)
library(plyr)
library(dplyr)
library(httr2)
library(knitr)
```

### Get APIs and keys from NYT section "Most Popular by Email" into R.

```{r API}
emailed_url <- fromJSON("https://api.nytimes.com/svc/mostpopular/v2/emailed/1.json?api-key=w8OnkqtBI6BsPbTAlCiAyQNJ0LCCRM7F", flatten = TRUE)
```

### Create a DafaFrame of 4 columns.

```{r DataFrame}
emailed_df <- ldply(emailed_url[4], data.frame)
```

### Check names within the new DataFrame.

```{r Names}
names(emailed_df)

```
### Select data to work with, and display it in console.

```{r select}
MostEmailed <- emailed_df %>% select(url, title, section, byline, published_date)
head(MostEmailed) 
```
```{r}
as.numeric(MostEmailed$published_date)
MostEmailed
```

### Final view of DataFrame.

```{r kable}
kable(MostEmailed)
```

### Analyze data, checked most emailed articles on NYTimes.com from 2020 by section.

```{r analyze}
MostEmailed %>% 
  group_by(section,published_date) %>%
  filter(published_date > '2020-01-01') %>%
  ggplot() +
  geom_bar(aes(x = section, fill = published_date)) +
  ggtitle("Most Emailed Articles") +
  xlab("Sections") +
  ylab("Count") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(fill = "Published Year")
```

## Conclussion: 

#### NY Times.com offers a platfiom for developers where we can access their APIs, after I created an account with them, I was able to obtain a key access to their website and from there I was able to get the "most emailed articles API" and import the Json file into R. After that I was able to create a DataFrame, select the data to work with, analyzed and visualized the section and published_data column.

