---
title: "Week10_Assignment"
author: "Victor Torres"
date: "2024-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction.

### For this Assignment, I'm going to get a primary example code from "Text Mining with R, Chapter 2"(Julia Silge and David Robinson.).Also, I will work with a different corpus of my choosing, and incorporate at one additional sentiment lexicon from another R package found through research.

## Load Libraries required for this assignment. 

```{r libraries}
library(janeaustenr)
library(tidyverse)
library(tidyr)
library(stringr)
library(tidytext)
library(dplyr)
library(stringr)
```

## Get specific sentiment lexicons from packages

```{r afinn}
get_sentiments("afinn")
```

```{r bing}
get_sentiments("bing")
```

```{r nrc}
get_sentiments("nrc")
```

## Sentiment Analysis with inner join.
```{r sentiment analysis with inner join}
tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```

## Filter DataFrame from book Emma.

```{r filter data}
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```

## Use pivot_wider() function to have a neg and pos sentiment in separate columns.

```{r use pivot_wider}
jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

## Plot sentiment scores

```{r plot1}
library(ggplot2)

ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")

```

## Comparison of the three sentiment dicctionaries.

```{r comparation}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice
```

##Use inner_join function to calculate sentiment in different ways.

```{r inner_join2}
afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

## Data visualization of net sentiment

```{r plot 2}
bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```

## Check how many positive and negative words are in lexicons

```{r p and n}
get_sentiments("nrc") %>% 
  filter(sentiment %in% c("positive", "negative")) %>% 
  count(sentiment)
get_sentiments("bing") %>% 
  count(sentiment)
```

## Use count() function to find out how much each word contribute to each sentiment.

```{r count}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()
```

## Visualization of word count

```{r plot 02}
bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```

## Use the bind_rows to spot certain words in the book.

```{r bind rows}
custom_stop_words <- bind_rows(tibble(word = c("miss"),  
                                      lexicon = c("custom")), 
                               stop_words)
```

## Wordcloud library

```{r wordcloud}
library(wordcloud)
tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```

## Wordclouds 2.

```{r wordclouds2}
library(reshape2)

tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)
```

## Tokenize text into sentences.

```{r token}
p_and_p_sentences <- tibble(text = prideprejudice) %>% 
  unnest_tokens(sentence, text, token = "sentences")
p_and_p_sentences$sentence[2]
```

## Split text of book by chapter

```{r split}
austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

austen_chapters %>% 
  group_by(book) %>% 
  summarise(chapters = n())
```

## Filter negative words by chapter of novel

```{r negative}
bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())

tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  slice_max(ratio, n = 1) %>% 
  ungroup()
```

## Working with a different Corpus.

### I will be using a library called "Gutengergr" to get a a "Peter Pan" book to analize it.

```{r new corpus}
library(gutenbergr)
?gutenbergr
PeterPan <- gutenberg_download(16) 
```

## Analize new corpus and convert text to tidy format using the inner_join. 

```{r new join}
tidy_Peter <- PeterPan %>%
  mutate(linenumber = row_number(), chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", ignore_case = TRUE)))) %>%
  unnest_tokens(word, text)
tidy_Peter

```

## Analize sentiments with NRC package.

```{r new nrc}
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_Peter %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```

## Get Net sentiment of the Novel.

```{r net sentiment}
afinn <- tidy_Peter %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  tidy_Peter %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

##Visualization of each chunck of the novel for each sentiment lexicon.

```{r plotting}
bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```

## Use loughran as a new sentiment lexicon

```{r new lexicon}
get_sentiments("loughran")
```

# Analyze data with new lexicon

```{r filter and get sentiments}
lexiconnew <- get_sentiments("loughran") %>% 
  filter(sentiment == "positive" | sentiment =="negative")

net_Peter <- tidy_Peter %>%
  inner_join(get_sentiments("bing")) %>%
  count(index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)
```

## Visualization of Net sentiments.

```{r net plot}
ggplot(net_Peter, aes(index, sentiment)) +
  geom_col(show.legend = FALSE) 
```

# Conclussion.

### I was able to work with the code displayed in chapter 2 of the book "Text Mining with R" by Julia Slinge and David Robinson. I worked with a new corpus called Perter Pan, and also I was able to work with a different sentiment lexicon called "Loughran", and analyzed the sentiments of the book with it.

# Sources:

#### Silge J & Robinson D. (2017).Text Mining with R: A Tidy Approach. https://www.tidytextmining.com/

#### Robinson, D.(2023, Nov 12).Download and process public domain works in the Project Gutenberg collection. https://cran.r-project.org/web/packages/gutenbergr/index.html

#### Barrie, J.(1991, July) The Project Gutenberg eBook of Peter Pan, by James M. Barrie. https://www.gutenberg.org/files/16/16-h/16-h.htm
