---
title: "project 2"
author: "Victor Torres"
date: "2024-03-02"
output: html_document
---
# Project 2 #
### The goal of this assignment is to give you practice in preparing different datasets for downstream analysis work. ###
#### I going tol get the data from a .CSV file that I created. ####
```{r get data from .CSV file}
college_data<-read.csv('https://raw.githubusercontent.com/vitugo23/DATA607/main/project2/grad_rates_data.csv')
dim(college_data)
head(college_data)
tail(college_data)
```

#### Load the libraries required for the project. ####
```{r}
library(tidyr)
library(dplyr)
library(stringr)
```

#### Name columns on data frame. #####
```{r}
colnames(college_data)[1] = "race"
colnames(college_data)[2] = "total_graduates"
colnames(college_data)[3] = "Percentage"
colnames(college_data)[4] = "5Y_Grad_rate"
colnames(college_data)[5] = "total1"
colnames(college_data)[6] = "assos_dg_rate"
colnames(college_data)[7] = "total2"
colnames(college_data)[8] = "bachelor_dg_rate"
colnames(college_data)[9] = "total3"
colnames(college_data)[10] = "Master_dg_rate"
colnames(college_data)[11] = "total4"
colnames(college_data)[12] = "PHD_dg_rate"
colnames(college_data)[13] = "total5"
total_graduates <- colnames(college_data[2:length(college_data)])
glimpse(college_data)
```
#### Remove tittle rows in data frame. ####
```{r}
tidycollege <- tail(college_data, -1) %>%
  select(-c(1))

head(college_data)
tidycollege$Item
```
#### Remove commas, and change values from string to numerical. ####
```{r}
college_data[total_graduates] <- college_data[total_graduates] %>% apply(MARGIN = 2, FUN = function(x) as.numeric(str_remove(x,",")))

print(total_graduates)
```
#### Change format of data to long format ####
```{r}
l_data <- college_data %>% pivot_longer(cols=total_graduates,names_to = "Comparison", values_to = "total_Comp")
print(l_data)
```

```{r}
l_data <- l_data %>% mutate(Race = ifelse(race == "", NA, race))

print(l_data)
```

```{r}
l_data <- l_data %>% fill(race)
print(l_data)

```

```{r}
l_data <- l_data %>% janitor::clean_names()

print(l_data)
```

```{r}
t_data <- l_data %>% spread(key=percentage, value=comparison)

print(t_data)

```

```{r}
t_data <- t_data %>% janitor::clean_names()

print(t_data)
```

```{r}
library(ggplot2)
ggplot(college_data) +
  geom_point(mapping = aes(x=race, y=total_graduates))
```



