---
title: "Tidyverse"
author: "Victor Torres"
date: "2024-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

For this assignment, we have to collaborate around a project in GitHub.Create an example using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.

I selected to work with a club soccer predictions dataset from the fivethirtyeight.com website, and I will use dplyr, tidyr, ggplot2, and kable libraries to tidy the data in this dataset.

```{r libraries}
library(tidyverse)
library(kableExtra)
```

### Import files into R

```{r import files into R}
soccer_data <- read.csv("https://raw.githubusercontent.com/vitugo23/DATA607/main/tidyverse_project/soccer-spi/spi_global_rankings_intl.csv")
head(soccer_data)
```

### Renaming columns with dplyr:: rename().

```{r change names}
soccer_data %>% 
  rename( 
    team = name,
    zone = confed,
    offense = off,
    defense = def
  )
soccer_data
```

### Display countries with highest scores using slice function in dplyr, and display results using a kable extra table

```{r}
top_10 <- soccer_data |>
slice_max(spi, n = 10)
kable(top_10)
```


### Remove NA's values from rows using the drop_na function from tidyr

```{r remove}
soccer_data <- soccer_data %>% drop_na()

print(soccer_data)
```

### Get the dimensions of dataframe using the dim function

```{r dim}
dim(soccer_data)
```

### Get the mean, median and quantile of the columns.

```{r}
soccer_data %>%
 summarise (Mean = mean(soccer_data$off), Median = median(soccer_data$def), Quant = quantile(soccer_data$spi))
```

### Use ggplot to visualize some stats,

```{r scatter}
plot(soccer_data$def, soccer_data$off, main="defense vs offense")
```

### Use ggplot to create a boxplot

```{r boxplot}
boxplot(off~def,data = soccer_data,
        main="Boxplot of soccer table",
        xlab="offense",ylab="defense")
```

## Conclussion
I was able to obtain the data from a fivethirtyeight.com and create a vignette to demostrate how to use a few capabilities of tidyverse packages, I use some functions of dplyr, tidyr, ggplot, and kable.


