---
title: "Tidyverse"
author: "Victor Torres"
date: "2024-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

For this assignment, we have to collaborate around a project in GitHub.Create an example using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.

I selected to work with a club soccer predictions dataset from the fivethirtyeight.com website, and I will use dplyr, tidyr, ggplot2, and kable libraries to tidy the data in this dataset.

```{r libraries}
library(tidyverse)
library(kableExtra)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
